Expressivity

### 2021-06-07 del/8_expressivity_pres.tex
DEL: Presentation: Final additions and fixes after presentation

\begin{frame}{Bisimulation: Countably Many Atoms}
	$(M^1,s^1)$ and $(M^2,s^2)$ are not bisimilar (\textbf{back} is not satisfied),\\
	but $(M^1,s^1)$ and $(M^2,s^2)$ satisfy the same formulas. \pause
	%\begin{small}
	\begin{proof}
		Induction on formulas
		\begin{itemize}
			\item $s^1,s^2$ satisfy the same properties (none)
			\item negation and conjuction are easy
		\end{itemize} \pause
		Suppose $K\phi$ a formula that could distinguish $(M^1,s^1)$ and $(M^2,s^2)$, i.e. $(M^1,s^1) \models K\phi$ and $(M^2,s^2) \not\models K\phi$ (opposite is impossible).
		\begin{itemize}
			\item $\phi$ is true in all $\mathbb{N}$ states and false in $\omega$
			\item $\phi$ is finite so let $n_{\max} = \max \{n|p_n \text{ occurs in } \phi\}$\\
				then $(M^2,n_{\max}+1) \models \phi$ iff $(M^2,\omega) \models \phi$
		\end{itemize} \pause
		Contradiction. Therefore $(M^1,s^1) \equiv_{\lang_K} (M^2,s^2)$
	\end{proof}
	%\end{small}
\end{frame}

@@

\begin{frame}{The $\lang_K(P)$ game}
	\begin{small}
	For models $M = \langle S,R,V \rangle, M^\prime = \langle S^\prime,R^\prime,V^\prime \rangle$ and states $s \in S, s^\prime \in S^\prime$, the $n$-round $\lang_K(P)$ \textit{game} between spoiler and duplicator on $(M,s)$ and $(M^\prime,s^\prime)$ is the following \pause
	\begin{itemize}
		\item If $n=0$ then spoiler wins if $s$ and $s^\prime$ differ in any atomic property in $P$, else duplicator wins \pause
		\item If $n>0$ then spoiler can play any of the following moves
			\begin{itemize}
				\item \textbf{forth-move}: choose $\alpha \in A, t \in S \text{ s.t. } (s,t) \in R_a$.\\
				Duplicator responds with state $t^\prime \in S^\prime$ s.t. $(s^\prime,t^\prime) \in R_a^\prime$
				\item \textbf{back-move}: choose $\alpha \in A, t^\prime \in S^\prime \text{ s.t. } (s^\prime,t^\prime) \in R_a^\prime$.\\
				Duplicator responds with state $t \in S$ s.t. $(s,t) \in R_a$
			\end{itemize}
			Output is $(t,t^\prime)$ \pause
		\item Game continues with the new output state
	\end{itemize} \pause
	If either player cannot perform an action, that player loses\\
	If the output states differ in their atomic properties in $P$, spoiler wins\\
	If spoiler hasn't won after $n$ rounds, duplicator wins
	\end{small}
\end{frame}

@@

\begin{frame}{The $\lang_K(P)$ game}
	\setbeamercolor{block title}{bg=}
	\setbeamercolor{block body}{bg=}
	\begin{block}{Proof. (cont.)}
		$(\leftarrow)$ Suppose $(M,s) \equiv_{\lang_K^{n+1}} (M^\prime,s^\prime)$, we need to describe the duplicator's winning strategy, i.e. show that for all $t: (s,t) \in R_a$ that spoiler selects, there is a $t^\prime: (s^\prime,t^\prime) \in R_a^\prime: (M,t) \equiv_{\lang_K^n} (M^\prime,t^\prime)$\\
		Then by the first induction hypothesis duplicator has a winning strategy for the remaining subgame \pause
		\\[8pt]
		Suppose there is no such $t^\prime$, so for all $t: (s^\prime,t^\prime)$ spoiler has a winning strategy, i.e. (from first induction hypothesis) there is a formula $\phi_{t^\prime}: d(\phi) \leq n$ where $(M^\prime,t^\prime) \models \phi_{t^\prime}, (M,s) \not\models \phi_{t^\prime}$
	\end{block}
\end{frame}

\begin{frame}{The $\lang_K(P)$ game}
	\setbeamercolor{block title}{bg=}
	\setbeamercolor{block body}{bg=}
	\begin{proof}[Proof. (cont.)]
		From Lemma 8.20 the set $\{\phi_{t^\prime} | (s^\prime,t^\prime) \in R_a^\prime\}$ is finite wrt equivalence.
		If $f$ a function that chooses one formula from each equivalence class $[t^\prime]$, then
		\[ \phi = \bigvee_{(t^\prime,u^\prime)\in R_{\alpha}^\prime} f([t^\prime]) \]
		is finite with $d(\phi) \leq n$ \pause
		\\[4pt]
		We have $(M^\prime,s^\prime) \models K_a \phi$ but $(M,s) \not\models K_a \phi$ while $d(K_a \phi) \leq n+1$\\
		Contradiction due to initial assumtion that $(M,s) \equiv_{\lang_K^{n+1}} (M^\prime,s^\prime)$\\
	\end{proof}
\end{frame}

\begin{frame}{The $\lang_K(P)$ game}
	\begin{block}{Theorem 8.22}
		$(M,s) \equiv (M^\prime,s^\prime)$ iff for all $n \in \mathbb{N}$ duplicator has a winning strategy for the $n$-round $\lang_K$-game on $(M,s)$ and $(M^\prime,s^\prime)$
	\end{block} \pause
	Games are finite but there are infinite games
\end{frame}

>>>>

\begin{frame}{Bisimulation: Countably Many Atoms}
	$(M^1,s^1)$ and $(M^2,s^2)$ are not bisimilar (\textbf{back} is not satisfied),\\
	but $(M^1,s^1)$ and $(M^2,s^2)$ satisfy the same formulas. \pause
	%\begin{small}
	\begin{proof}
		Induction on formulas
		\begin{itemize}
			\item $s^1,s^2$ satisfy the same atoms (none)
			\item negation and conjuction are easy
		\end{itemize} \pause
		Suppose $K\phi$ a formula that could distinguish $(M^1,s^1)$ and $(M^2,s^2)$, i.e. $(M^1,s^1) \models K\phi$ and $(M^2,s^2) \not\models K\phi$ (opposite is impossible).
		\begin{itemize}
			\item $\phi$ is true in all $\mathbb{N}$ states and false in $\omega$
			\item $\phi$ is finite so let $n_{\max} = \max \{n|p_n \text{ occurs in } \phi\}$\\
				then $(M^2,n_{\max}+1) \models \phi$ iff $(M^2,\omega) \models \phi$
		\end{itemize} \pause
		Contradiction. Therefore $(M^1,s^1) \equiv_{\lang_K} (M^2,s^2)$
	\end{proof}
	%\end{small}
\end{frame}

@@

\begin{frame}{The $\lang_K(P)$ game}
	\begin{small}
	For models $M = \langle S,R,V \rangle, M^\prime = \langle S^\prime,R^\prime,V^\prime \rangle$ and states $s \in S, s^\prime \in S^\prime$, the $n$-round $\lang_K(P)$ \textit{game} between spoiler and duplicator on $(M,s)$ and $(M^\prime,s^\prime)$ is the following \pause
	\begin{itemize}
		\item If $n=0$ then spoiler wins if $s$ and $s^\prime$ differ in any atomic property in $P$, else duplicator wins \pause
		\item If $n>0$ then spoiler can play any of the following moves
			\begin{itemize}
				\item \textbf{forth-move}: choose $a \in A, t \in S \text{ s.t. } (s,t) \in R_a$.\\
				Duplicator responds with state $t^\prime \in S^\prime$ s.t. $(s^\prime,t^\prime) \in R_a^\prime$
				\item \textbf{back-move}: choose $a \in A, t^\prime \in S^\prime \text{ s.t. } (s^\prime,t^\prime) \in R_a^\prime$.\\
				Duplicator responds with state $t \in S$ s.t. $(s,t) \in R_a$
			\end{itemize}
			Output is $(t,t^\prime)$ \pause
		\item Game continues with the new output state
	\end{itemize} \pause
	If either player cannot perform an action, that player loses\\
	If the output states differ in their atomic properties in $P$, spoiler wins\\
	If spoiler hasn't won after $n$ rounds, duplicator wins
	\end{small}
\end{frame}

@@

\begin{frame}{The $\lang_K(P)$ game}
	\setbeamercolor{block title}{bg=}
	\setbeamercolor{block body}{bg=}
	\begin{block}{Proof. (cont.)}
		$(\leftarrow)$ Suppose $(M,s) \equiv_{\lang_K^{n+1}} (M^\prime,s^\prime)$, we need to describe the duplicator's winning strategy, i.e. show that for all $t: (s,t) \in R_a$ that spoiler selects, there is a $t^\prime: (s^\prime,t^\prime) \in R_a^\prime: (M,t) \equiv_{\lang_K^n} (M^\prime,t^\prime)$\\
		Then by the first induction hypothesis duplicator has a winning strategy for the remaining subgame \pause
		\\[8pt]
		Suppose there is no such $t^\prime$, so for all $t^\prime: (s^\prime,t^\prime)$ spoiler has a winning strategy, i.e. (from first induction hypothesis) there is a formula $\phi_{t^\prime}: d(\phi) \leq n$ where $(M^\prime,t^\prime) \models \phi_{t^\prime}, (M,t) \not\models \phi_{t^\prime}$
	\end{block}
\end{frame}

\begin{frame}{The $\lang_K(P)$ game}
	\setbeamercolor{block title}{bg=}
	\setbeamercolor{block body}{bg=}
	\begin{proof}[Proof. (cont.)]
		From Lemma 8.20 the set $\{\phi_{t^\prime} | (s^\prime,t^\prime) \in R_a^\prime\}$ is finite wrt equivalence.
		If $f$ a function that chooses one formula from each equivalence class $[t^\prime]$, then
		\[ \phi = \bigvee_{(s^\prime,t^\prime)\in R_{\alpha}^\prime} f([t^\prime]) \]
		is finite with $d(\phi) \leq n$ \pause
		\\[4pt]
		We have $(M^\prime,s^\prime) \models K_a \phi$ but $(M,s) \not\models K_a \phi$ while $d(K_a \phi) \leq n+1$\\
		Contradiction due to initial assumtion that $(M,s) \equiv_{\lang_K^{n+1}} (M^\prime,s^\prime)$\\
	\end{proof}
\end{frame}

\begin{frame}{The $\lang_K(P)$ game}
	\begin{block}{Theorem 8.22}
		$(M,s) \equiv (M^\prime,s^\prime)$ iff for all $n \in \mathbb{N}$ duplicator has a winning strategy for the $n$-round $\lang_K$-game on $(M,s)$ and $(M^\prime,s^\prime)$
	\end{block} \pause
	Games are finite but there are infinite games\\ \pause
	Winning strategy for duplicator coincides with bisimulation only when we consider an $\lang_K(P)$ game with $\omega$ rounds (the first ordinal limit)
\end{frame}
